Automated inference of production rules for glycans

### 2021-07-05 papers/sugar-synth/algo.tex
dsadsa

In this section, we present a method to solve the synthesis problem of finding
a set of production rules that produce a given set $\mu$ of molecules.
Our synthesis method \textsc{SugarSynth} is presented in Algorithm~\ref{alg:sugar-synth}.
Here, we are considering only the single compartment case.
The generalizations are discussed in
\choosefinal{extended version~\cite{arxiv-this}.}{ Appendix~\ref{sec:variations}.}

@@

\begin{theorem}
  If \textsc{SugarSynth}($\mu$, $d$, $n$) returns a set of rules $R$, then if $R$ produces a molecule that has depth less than $h$, the it is in $\mu$ (soundness).
  If it fails to synthesize rules, there are no rule set from the budget of depth $d$ and number $n$ (completeness).
\end{theorem}

>>>>

In this section, we present a method to solve the synthesis problem of finding
a set of production rules that produce a given set $\mu$ of molecules.
Our synthesis method \textsc{SugarSynth} is presented in Algorithm~\ref{alg:sugar-synth}.
Here, we are considering only the single compartment case.
The generalizations are discussed in
\choosefinal{the extended version~\cite{arxiv-this}.}{ Appendix~\ref{sec:variations}.}

@@

\begin{theorem}
  If \textsc{SugarSynth}($\mu$, $d$, $n$) returns a set of rules $R$, then if $R$ produces a molecule that has depth less than $h+1$, the it is in $\mu$ (soundness).
  If it fails to synthesize rules, there are no rule set from the budget of depth $d$ and number $n$ (completeness).
\end{theorem}

### 2021-07-05 papers/sugar-synth/intro.tex
dsadsa

In this paper, we are presenting the {\em first} automated synthesis
method for the production rules.
Our method takes the observed glycan molecules in a cell as input and synthesizes
the possible production rules that may explain the observation.
Our work is the {\em first} to consider the computational problem.
Our method of synthesis is similar to counterexample guided
inductive synthesis(CEGIS)~\cite{cegis}.
Several methods for solving problems of searching in a complex combinatorial
space use templates to define and limit their search space,
such as learning invariants of programs~\cite{InvGenTACAS09},
and synthesizing missing components in programs~\cite{sygus,Solar-Lezama2005}.
We also use templates to model the production rules.
% The templates usually have parameters that limit the search space.
% In our case, the size and number of templates are the parameters that
% limit sizes of pieces of trees that are added in one step,
% number of the production rules, and
% the depth of surrounding that are checked to control the rules.

@@

We append our synthesis constraints with additional constraints stating that
no matter how we apply the synthesized rules, they will not produce the extra molecule.
Since there is a requirement that {\em all} possible applications of rules must satisfy
a condition, we have quantifiers in the constraints.
We use a solver that handles quantifiers over finite range variable in
the synthesis query.
We go to the next iteration of the method.
The method always terminates because the search space of rules is finite. The set of rules synthesized need not minimal. They also need not be unique. The solver reports the first set which satisfies the constraints.

Our encoding to constraints depends on the model of execution of the rules.
The current biological information is not sufficient to make a precise
and definite model, and do the synthesis.
We have also explored the variations of the models. For example, all rules
may apply simultaneously. They apply in batches because they
stay in different compartments.
The molecule under assembly may flow through the compartments.
The distribution of the stay of the molecules in a compartment also affects
the execution model.
Furthermore, we may have variations in the type and quality of data available to us.
\choosefinal{We support the variations, which is presented in 
  the extended version of this paper is available at~\cite{arxiv-this}.}{We support the variations.}

@@

We organize the paper as follows.
In Section~\ref{sec:bio}, we introduce the biological background.
%of the synthesis problem and present the basis of the formal model.
% If the reader wants to avoid getting into the biological details, she may skip the section.
In Section~\ref{sec:motivation}, we present a motivating example to illustrate our method.
In Section~\ref{sec:model}, we present the formal model of the glycans
and their production rules.
In Section~\ref{sec:algo}, we present our method for the synthesis problem.
%and the modifications in our methods to support the variations.
In Section~\ref{sec:experiments}, we present our implementation  and experiments. %on some glycan data.
In Section~\ref{sec:conclusion}, we conclude the paper.
\choosefinal{}{We present the extended related work in Appendix~\ref{sec:related}.}

>>>>

In this paper, we are presenting the {\em first} automated synthesis
method for the production rules.
Our method takes the observed glycan molecules in a cell as input and synthesizes
the possible production rules that may explain the observation.
To our knowledge,
our work is the {\em first} to consider the computational problem.
Our method of synthesis is similar to counterexample guided
inductive synthesis(CEGIS)~\cite{cegis}.
Several methods for solving problems of searching in a complex combinatorial
space use templates to define and limit their search space,
such as learning invariants of programs~\cite{InvGenTACAS09},
and synthesizing missing components in programs~\cite{sygus,Solar-Lezama2005}.
We also use templates to model the production rules.
% The templates usually have parameters that limit the search space.
% In our case, the size and number of templates are the parameters that
% limit sizes of pieces of trees that are added in one step,
% number of the production rules, and
% the depth of surrounding that are checked to control the rules.

@@

We append our synthesis constraints with additional constraints stating that
no matter how we apply the synthesized rules, they will not produce the extra molecule.
Since there is a requirement that {\em all} possible applications of rules must satisfy
a condition, we have quantifiers in the constraints.
We use a solver that handles quantifiers over finite range variable in
the synthesis query.
We go to the next iteration of the method.
The method always terminates because the search space of rules is finite.
The set of rules synthesized need not be minimal or unique. The solver reports
the first set which satisfies the constraints.
However, our method is adoptable. We can add various optimization criteria to
find optimal rules for the given objectives, e.g., smallest rule sizes,
number of rules, etc.

Our encoding to constraints depends on the model of execution of the rules.
The current biological information is not sufficient to make a precise
and definite model, and do the synthesis.
We have also explored the variations of the models. For example, all rules
may apply simultaneously. They apply in batches because they
stay in different compartments.
The molecule under assembly may flow through the compartments.
The distribution of the stay of the molecules in a compartment also affects
the execution model.
Furthermore, we may have variations in the type and quality of data available to us. For example, we may have missed a produced molecule in experiments.
\choosefinal{We support the variations, which is presented in 
  the extended version of this paper~\cite{arxiv-this}.}{We support the variations.}

@@

We organize the paper as follows.
In Section~\ref{sec:bio}, we introduce the biological background.
%of the synthesis problem and present the basis of the formal model.
% If the reader wants to avoid getting into the biological details, she may skip the section.
In Section~\ref{sec:motivation}, we present a motivating example to illustrate our method.
In Section~\ref{sec:model}, we present the formal model of the glycans
and their production rules.
In Section~\ref{sec:algo}, we present our method for the synthesis problem.
%and the modifications in our methods to support the variations.
In Section~\ref{sec:experiments}-~\ref{sec:conclusion}, we present our experiments and conclude the paper. %on some glycan data.
% In Section~\ref{sec:conclusion}, we conclude the paper.
\choosefinal{}{We present the extended related work in Appendix~\ref{sec:related}.}
